---
title: "R Assignment"
author: "Dylan Harris"
date: "2023-03-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data Inspection
```{r}
library(tidyverse)
genotypes <- read_tsv("https://github.com/dah96/bcb546/raw/master/R_Assignment/fang_et_al_genotypes.txt")
snp_positions <- read_tsv("https://github.com/dah96/bcb546/raw/master/R_Assignment/snp_position.txt")
```
Loaded the tidyverse package and loaded the two data files as dataframes in R.
```{r}
nrow(genotypes)
ncol(genotypes)
object.size(genotypes)
```
fang_et_al_genotypes.txt has 2782 rows, 986 columns, and is 23.1Mb in size.
```{r}
nrow(snp_positions)
ncol(snp_positions)
object.size(snp_positions)
```
snp_position.txt is 983 rows, 15 columns, and is 359kb is size.

# Data Processing
```{r}
maize_genotypes <- filter(genotypes, Group %in% c("ZMMIL","ZMMLR","ZMMMR"))
teosinte_genotypes <- filter(genotypes, Group %in% c("ZMPBA","ZMPJA","ZMPIL"))
```
Filtered genotypes into maize and teosinte genotypes respectively.
```{r}
t_m_genotypes <- t(maize_genotypes) %>%
  data.frame()
colnames(t_m_genotypes) <- t_m_genotypes[c(1),]
t_m_genotypes <- t_m_genotypes[-c(1,2,3),]
t_t_genotypes <-  t(teosinte_genotypes) %>%
  data.frame()
colnames(t_t_genotypes) <- t_t_genotypes[c(1),]
t_t_genotypes <- t_t_genotypes[-c(1,2,3),]
```
Transposes the maize genotypes, but doing so converts the dataframe into a matrix, which we turn back into a dataframe. The column names also need to be adjusted to the first row of the transposed data frame, then the superfluous rows are removed. Same process is repeated for teosinte genotypes.
```{r}
snp_cut <- select(snp_positions, SNP_ID, Chromosome, Position)
maize_joined <- bind_cols(snp_cut, t_m_genotypes)
teosinte_joined <-  bind_cols(snp_cut, t_t_genotypes)
```
snp_positions is used to create a new dataframe only containing SNP_ID, Chromosome, and Positions columns which is then joined with the transposed maize and teosinte genotypes respectively.
```{r}
maize_chr <- vector(length = 10)
for (i in 1:10) {
  maize_chr[[i]] <- filter(maize_joined, Chromosome == i) %>%
    arrange(as.numeric(Position))
}
maize_chr[[1]] <- filter(maize_joined, Chromosome == 1) %>%
  arrange(as.numeric(Position))
maize_chr2 <- filter(maize_joined, Chromosome == 2) %>%
  arrange(as.numeric(Position))
maize_chr3 <- filter(maize_joined, Chromosome == 3) %>%
  arrange(as.numeric(Position))
maize_chr4 <- filter(maize_joined, Chromosome == 4) %>%
  arrange(as.numeric(Position))
maize_chr5 <- filter(maize_joined, Chromosome == 5) %>%
  arrange(as.numeric(Position))
maize_chr6 <- filter(maize_joined, Chromosome == 6) %>%
  arrange(as.numeric(Position))
maize_chr7 <- filter(maize_joined, Chromosome == 7) %>%
  arrange(as.numeric(Position))
maize_chr8 <- filter(maize_joined, Chromosome == 8) %>%
  arrange(as.numeric(Position))
maize_chr9 <- filter(maize_joined, Chromosome == 9) %>%
  arrange(as.numeric(Position))
maize_chr10 <- filter(maize_joined, Chromosome == 10) %>%
  arrange(as.numeric(Position))
```
Maize chromosomes pulled from joined dataframe via filter and arranged via numerical position.
```{r}
maize_alt_joined <- data.frame(lapply(maize_joined, function(x) {
  gsub("[?]", "-", x)
}))
maize_alt_chr1 <- filter(maize_alt_joined, Chromosome == 1) %>% arrange(desc(as.numeric(Position)))
maize_alt_chr2 <- filter(maize_alt_joined, Chromosome == 2) %>% arrange(desc(as.numeric(Position)))
maize_alt_chr3 <- filter(maize_alt_joined, Chromosome == 3) %>% arrange(desc(as.numeric(Position)))
maize_alt_chr4 <- filter(maize_alt_joined, Chromosome == 4) %>% arrange(desc(as.numeric(Position)))
maize_alt_chr5 <- filter(maize_alt_joined, Chromosome == 5) %>% arrange(desc(as.numeric(Position)))
maize_alt_chr6 <- filter(maize_alt_joined, Chromosome == 6) %>% arrange(desc(as.numeric(Position)))
maize_alt_chr7 <- filter(maize_alt_joined, Chromosome == 7) %>% arrange(desc(as.numeric(Position)))
maize_alt_chr8 <- filter(maize_alt_joined, Chromosome == 8) %>% arrange(desc(as.numeric(Position)))
maize_alt_chr9 <- filter(maize_alt_joined, Chromosome == 9) %>% arrange(desc(as.numeric(Position)))
maize_alt_chr10 <- filter(maize_alt_joined, Chromosome == 10) %>% arrange(desc(as.numeric(Position)))
```

```{r}
teosinte_chr1 <- filter(teosinte_joined, Chromosome == 1) %>%
  arrange(as.numeric(Position))
teosinte_chr2 <- filter(teosinte_joined, Chromosome == 2) %>%
  arrange(as.numeric(Position))
teosinte_chr3 <- filter(teosinte_joined, Chromosome == 3) %>%
  arrange(as.numeric(Position))
teosinte_chr4 <- filter(teosinte_joined, Chromosome == 4) %>%
  arrange(as.numeric(Position))
teosinte_chr5 <- filter(teosinte_joined, Chromosome == 5) %>%
  arrange(as.numeric(Position))
teosinte_chr6 <- filter(teosinte_joined, Chromosome == 6) %>%
  arrange(as.numeric(Position))
teosinte_chr7 <- filter(teosinte_joined, Chromosome == 7) %>%
  arrange(as.numeric(Position))
teosinte_chr8 <- filter(teosinte_joined, Chromosome == 8) %>%
  arrange(as.numeric(Position))
teosinte_chr9 <- filter(teosinte_joined, Chromosome == 9) %>%
  arrange(as.numeric(Position))
teosinte_chr10 <- filter(teosinte_joined, Chromosome == 10) %>%
  arrange(as.numeric(Position))
```
```{r}
teosinite_alt_joined <- data.frame(lapply(teosinte_joined, function(x) {
  gsub("[?]", "-", x)
}))
teosinte_chr10 <- filter(teosinte_joined, Chromosome == 10) %>%
  arrange(as.numeric(Position))
```

# Data Visualization

```{r}

```

